---
title: "Estimating parameters"
author: "Olga Usuga and Freddy Hernandez"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "##"
)
```

# Introduction

In this vignette we explain how to use the functions available in the package to estimate parameters for each new distribution. First we need to require the __RelDists__ package.

```{r}
require(RelDists)
```

## Exponentiated Weibull Distribution
First, we simulate 100 values for the distribution using known parameters.

```{r}
y <- rEW(n=100, mu=2, sigma=1.5, nu=0.5)
```

Now we use the `gamlss` function from __gamlss__ package to estimate the parameters.

```{r}
# Fitting the model
mod <- gamlss::gamlss(y~1, sigma.fo=~1, nu.fo=~1, family='EW')
```

To obtain the fitted values for $\mu$, $\sigma$ and $\nu$ we use the inverse link function.

```{r}
exp(coef(mod, what='mu'))
exp(coef(mod, what='sigma'))
exp(coef(mod, what='nu'))
```



