---
title: "EW distribution"
author: "Olga Usuga and Freddy Hernandez"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "##"
)
```

# Exponentiated Weibull distribution

In this vignette we explore the family `EW` family to estimate parameters. The distribution was proposed by Mudholkar and Srivastava (1993).
.
```{r, message=FALSE}
require(RelDists)
```

## 1) Without covariates

First, we simulate 100 values for the distribution using known parameters.

```{r}
y <- rEW(n=100, mu=2.0, sigma=1.5, nu=0.5)
```

Next figure shows the density for the `y` vector.

```{r density01, echo=FALSE}
plot(density(y), lwd=2, col='blue3', main='')
rug(y, col='blue3')
```

Now we use the `gamlss` function from __gamlss__ package to estimate the parameters.

```{r, message=FALSE}
require(gamlss)
mod <- gamlss(y~1, sigma.fo=~1, nu.fo=~1, family='EW',
              control=gamlss.control(n.cyc=250, trace=FALSE))
```

To obtain the fitted values for $\mu$, $\sigma$ and $\nu$ we use the inverse link function $\log()$.

```{r}
exp(coef(mod, what='mu'))
exp(coef(mod, what='sigma'))
exp(coef(mod, what='nu'))
```

The results show that estimated parameters are close the true parameters $\mu$, $\sigma$ and $\nu$.

## 2) With covariates
Here ...
