---
title: "EW distribution"
author: "Olga Usuga and Freddy Hernandez"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "##"
)
```

# Exponentiated Weibull distribution

In this vignette we explore the family `EW` family to estimate parameters. The distribution was proposed by Mudholkar and Srivastava (1993).
.
```{r, message=FALSE}
require(RelDists)
```

## 1) Without covariates

### Example from Mudholkar and Srivastava (1993) page 301

```{r density01}
y <- c(0.1, 0.2, 1, 1, 1, 1, 1, 2, 3, 6, 7, 11, 12, 18, 18, 18, 18, 18, 21,
       32, 36, 40, 45, 46, 47, 50, 55, 60, 63, 63, 67, 67, 67, 67, 72, 75, 
       79, 82, 82, 83, 84, 84, 84, 85, 85, 85, 85, 85, 86, 86)

plot(density(y), lwd=2, col='blue3', main='')
rug(y, col='blue3')
```

Unfortunately the next code does not work.

```{r, eval=FALSE}
require(gamlss)
mod <- gamlss(y~1, sigma.fo=~1, nu.fo=~1, family='EW',
              control=gamlss.control(n.cyc=100, trace=FALSE))
```

### Example from Mudholkar and Hutson (1996) page 3077

```{r density02}
y <- c(1460, 4050, 3570, 2060, 1300, 1390, 1720, 6280, 1360, 7440,
       5320, 1400, 3240, 2710, 4520, 4840, 8320, 13900, 71500, 6250,
       2260, 318, 1330, 970, 1920, 15100, 2870, 20600, 3810, 726,
       7500, 7170, 2000, 829, 17300, 4740, 13400, 1940, 5660)
plot(density(y), lwd=2, col='blue3', main='')
```


### Using simulated data
First, we simulate 100 values for the distribution using known parameters.

```{r}
y <- rEW(n=100, mu=2.0, sigma=1.5, nu=0.5)
```

Next figure shows the density for the `y` vector.

```{r density03, echo=FALSE}
plot(density(y), lwd=2, col='blue3', main='')
rug(y, col='blue3')
```

Now we use the `gamlss` function from __gamlss__ package to estimate the parameters.

```{r, message=FALSE}
require(gamlss)
mod <- gamlss(y~1, sigma.fo=~1, nu.fo=~1, family='EW',
              control=gamlss.control(n.cyc=250, trace=FALSE))
```

To obtain the fitted values for $\mu$, $\sigma$ and $\nu$ we use the inverse link function $\log()$.

```{r}
exp(coef(mod, what='mu'))
exp(coef(mod, what='sigma'))
exp(coef(mod, what='nu'))
```

The results show that estimated parameters are close the true parameters $\mu$, $\sigma$ and $\nu$.



## 2) With covariates
Here ...
