---
title: "EW distribution"
author: "Olga Usuga and Freddy Hernandez"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "##"
)
```

## Exponentiated Weibull distribution

In this vignette we explore the Exponentiated Weibull distribution.

The distribution was proposed by Mudholkar and Srivastava (1993) but in this document we used the parameterization given by Almalki and Nadarajah (2014).

The pdf of the EW distribution is:

$$f(x) = \nu \mu \sigma x^{\sigma-1} \exp(-\mu x^\sigma)
      (1-\exp(-\mu x^\sigma))^{\nu-1}, \quad x>0,$$
      
where $\mu$, $\sigma$, $\nu$ > 0; $\mu$ is a scale parameter and $\sigma$ and $\nu$ are shape parameters.

For some values of the parameters, the Exponentiated Weibull distribution yiels some known distributions, the next table shows those cases.

|$\mu$      |  $\nu$      |   $\sigma$  |     *Distribution*               |
|:----------|:----------: | :---------: |   ----------------------------   |
|      1    |             |             |      Weibull                     | 
|           |             |        2    |      Generalized Rayleigh (GR)   |
|           |             |        1    |      Generalized Exponential(GE) |
|           |      1      |        1    |      Exponential                 |
|           |      1      |        2    |      Rayleigh                    |
 

The next figure shows possible shapes of the probability density function for several values of the parameters:

```{r, pdfEW1, echo=FALSE, message=F}
require(RelDists)
curve(dEW(x, 5, 0.5, 10), from=0, to=2, las=1, xlab='Density',
      ylim=c(0,3), col="red", main="Probability density function", las=1)
curve(dEW(x, 0.7, 0.5, 0.15), col="green3", add=T)
curve(dEW(x, 1, 3, 1), col="purple", add=T)

cap1 <- expression(paste(mu, "=", 5, ", ", sigma, "=", 0.5, ", ", nu, "=", 10))
cap2 <- expression(paste(mu, "=", 5, ", ", sigma, "=", 0.5, ", ", nu, "=", 10))
cap3 <- expression(paste(mu, "=", 5, ", ", sigma, "=", 0.5, ", ", nu, "=", 10))

legend(x=1.4, y=3.0, 
       legend=c(cap1, cap2, cap3), 
       col=c("red", "green3", "purple"), lty=c(1, 1, 1), bty="n")
```   


## 1) Estimation parameters without covariates

### Example from Mudholkar and Hutson (1996) page 3077
The Floyd River flood rates are in vector `y` and the objective is to find the parameter for the exponentiated Weibull distribution. 

```{r density02}
y <- c(1460, 4050, 3570, 2060, 1300, 1390, 1720, 6280, 1360, 7440,
       5320, 1400, 3240, 2710, 4520, 4840, 8320, 13900, 71500, 6250,
       2260, 318, 1330, 970, 1920, 15100, 2870, 20600, 3810, 726,
       7500, 7170, 2000, 829, 17300, 4740, 13400, 1940, 5660)
```

To estimate the parameter for the exponentiated Weibull distribution we use the next code.

```{r message=FALSE}
require(gamlss)
mod <- gamlss(y~1, sigma.fo=~1, nu.fo=~1, family='EW',
              control=gamlss.control(n.cyc=1500, trace=FALSE))
```

The results for the fitted model can be found below.

```{r}
exp(coef(mod, what='mu'))
exp(coef(mod, what='sigma'))
exp(coef(mod, what='nu'))
```

In the next figure we have the empirical density (blue) and the fitted density (black). We note that the fitted distribution explains better the observed data.

```{r pdfEW2}
plot(density(y), lwd=2, col='blue3', main='', las=1, ylim=c(0, 0.0002))
curve(dEW(x, mu=0.23, sigma=0.33, nu=19.03), 
      from=0.01, add=TRUE, lwd=2)
```

### Using simulated data
First, we simulate 100 values for the distribution using known parameters.

```{r}
y <- rEW(n=100, mu=5, sigma=0.5, nu=10)
```

Now we use the `gamlss` function from __gamlss__ package to estimate the parameters.

```{r, message=FALSE}
mod <- gamlss(y~1, sigma.fo=~1, nu.fo=~1, family='EW',
              control=gamlss.control(n.cyc=250, trace=FALSE))
```

To obtain the fitted values for $\mu$, $\sigma$ and $\nu$ we use the inverse link function $\log()$.

```{r}
exp(coef(mod, what='mu'))
exp(coef(mod, what='sigma'))
exp(coef(mod, what='nu'))
```

The results show that estimated parameters are close to the true parameters $\mu$, $\sigma$ and $\nu$.

In the next figure we have the empirical density (blue) and the fitted density (black). We note that the fitted distribution explains better the observed data.

```{r pdfEW3}
plot(density(y), lwd=2, col='blue3', main='', las=1)
curve(dEW(x, mu=exp(coef(mod, what='mu')), 
          sigma=exp(coef(mod, what='sigma')), 
          nu=exp(coef(mod, what='nu'))), 
      from=0.01, add=TRUE, lwd=2)
```

### Example from Mudholkar et al. (1995) page 443
The data set corresponds to survival times of 51 head-and-neckcancer patients. The times are in days but they are converted to months (1 month = 30.438 days) as in the original paper. Below are the data.

```{r pdfEW4}
dt <- data.frame(y=c(7, 34, 42, 63, 64, 74, 83, 84, 91, 108, 112, 129, 
                     133, 133, 139, 140, 140, 146, 149, 154, 157, 160,
                     160, 165, 173, 176, 185, 218, 225, 241, 248, 273, 
                     277, 279, 297, 319, 405, 417, 420, 440, 523, 523, 
                     583, 594, 1101, 1116, 1146, 1226, 1349, 1412, 
                     1417) / 30.438, 
                 status=c(TRUE, TRUE, TRUE, TRUE, TRUE, FALSE, TRUE, 
                          TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, 
                          TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, 
                          TRUE, TRUE, TRUE, TRUE, TRUE, FALSE, TRUE, 
                          TRUE, TRUE, TRUE, TRUE, TRUE, FALSE, TRUE, 
                          FALSE, TRUE, TRUE, TRUE, TRUE, TRUE, FALSE, 
                          TRUE, TRUE, TRUE, FALSE, TRUE, FALSE, 
                          FALSE, FALSE, TRUE))

plot(density(dt$y), main='')
rug(dt$y)
```

The variable `y` corresponds to survival time and `status` is a logical variable indicating if the patient died (`TRUE`) or not. To ensure convergence in the fit procedure we are going to use only survival times $< 40$ months.

```{r}
require(survival)
require(gamlss.cens)

mod <- gamlss(Surv(y, status) ~ 1,
              family=cens(EW), data=subset(dt, y < 40),
              control=gamlss.control(n.cyc=5000, trace=FALSE))
```

The fitted values are:

```{r}
exp(coef(mod, what='mu'))
exp(coef(mod, what='sigma'))
exp(coef(mod, what='nu'))
```

The hazard function for this example can be obtained using the next code and the figure coincides with the figure 4 in Mudholkar et al (1995).

```{r hazard1}
curve(hEW(x,
          mu=exp(coef(mod, what='mu')), 
          sigma=exp(coef(mod, what='sigma')), 
          nu=exp(coef(mod, what='nu'))),
      ylab='Hazard', xlab='Survival times (months)',
      from=0, to=50, las=1)
```



## 2) Estimation parameters with covariates

