---
title: "FWE distribution"
author: "Olga Usuga and Freddy Hernandez"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "##"
)
```

## Flexible Weibull extension distribution

In this vignette we explore the Flexible Weibull extension distribution.

The distribution was proposed by Bebbington (2007).

The pdf of the EW distribution is:
 
$$f(x) = \left( \mu+ \frac{\sigma}{x^2} \right) e^{\mu x - \sigma / x} \exp \left( -e^{\mu x - \sigma / x} \right), \quad x > 0,$$   
where $\mu > 0$ and $\sigma > 0$.

Next figure shows possible shapes of the probability density function for several values of the parameters:

```{r, pdfFWE1, echo=FALSE, message=F}
require(RelDists)
curve(dFWE(x, mu=0.75, sigma=0.5), from=0, to=3, ylim=c(0, 1.7), 
      col="red",ylab="Probability density function", las=1)
curve(dFWE(x, mu=2.00, sigma=3.0), col="blue", add=TRUE)
curve(dFWE(x, mu=0.75, sigma=1.3), col="purple", add=TRUE)

cap1 <- expression(paste(mu, "=", 0.75, ", ", sigma, "=", 0.5))
cap2 <- expression(paste(mu, "=", 2.00, ", ", sigma, "=", 3.0))
cap3 <- expression(paste(mu, "=", 0.75, ", ", sigma, "=", 1.3))

legend(x=2.0, y=1.8, 
       legend=c(cap1, cap2, cap3), 
       col=c("red", "blue", "purple"), lty=c(1, 1, 1), bty="n")
```  

## 1) Estimation parameters without covariates

### Example from Bebbington, Chin-Diew and Zitikis (2007) page 723
The data below correspond to the time between failures of secondary reactor pumps.

```{r}
y <- c(2.160, 0.746, 0.402, 0.954, 0.491, 6.560, 4.992, 0.347,
       0.150, 0.358, 0.101, 1.359, 3.465, 1.060, 0.614, 1.921,
       4.082, 0.199, 0.605, 0.273, 0.070, 0.062, 5.320)
```

To estimate the parameter for the Flexible Weibull extension distribution we use the next code.

```{r message=FALSE}
require(RelDists)
require(gamlss)
mod <- gamlss(y~1, sigma.fo=~1, family='FWE',
              control=gamlss.control(n.cyc=5000, trace=FALSE))
```

The results for the fitted model can be found below.

```{r}
exp(coef(mod, what='mu'))
exp(coef(mod, what='sigma'))
```

In the next figure we have the histogram for the data and the fitted density. We note that the fitted distribution explains better the observed data.

```{r fwe1}
hist(y, freq=FALSE, breaks=10, ylim=c(0, 1.7), las=1, main='')
curve(dFWE(x, mu=0.2065115, sigma=0.2588857), 
      from=0.01, add=TRUE, lwd=2)
```

### Using simulated data
First, we simulate 100 values for the distribution using known parameters.

```{r}
y <- rFWE(n=100, mu=0.75, sigma=0.5)
```

Now we use the `gamlss` function from __gamlss__ package to estimate the parameters.

```{r, message=FALSE}
mod <- gamlss(y~1, sigma.fo=~1, family='FWE',
              control=gamlss.control(n.cyc=250, trace=FALSE))
```

To obtain the fitted values for $\mu$, $\sigma$ and $\nu$ we use the inverse link function $\log()$.

```{r}
exp(coef(mod, what='mu'))
exp(coef(mod, what='sigma'))
```

The results show that estimated parameters are close to the true parameters $\mu$, $\sigma$ and $\nu$.

In the next figure we have the histogram for the data and the fitted density. We note that the fitted density explains better the observed data.

```{r fwe2}
hist(y, freq=FALSE, breaks=10, ylim=c(0, 1.5), las=1, main='')
curve(dFWE(x, mu=exp(coef(mod, what='mu')), 
          sigma=exp(coef(mod, what='sigma'))), 
      from=0.01, add=TRUE, lwd=2)
```



## 2) Estimation parameters with covariates

