---
title: "OW distribution"
author: "Jaime Mosquera"
date: "`r Sys.Date()`"
bibliography: ../inst/REFERENCES.bib
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{OW distribution}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(RelDists)
```

# Odd Weibull distribution

This distribution was proposed by Cooray [-@Cooray2006]. The probability density function $f(x)$ and cumulative density function $F(x)$ are given by:

$$f(x) = \left( \frac{\sigma\nu}{x} \right) (\mu x)^\sigma e^{(\mu x)^\sigma} \left(e^{(\mu x)^{\sigma}}-1\right)^{\nu-1} \left[ 1 + \left(e^{(\mu x)^{\sigma}}-1\right)^\nu \right]^{-2},$$

and

$$F(x) = 1 - \left[ 1 + \left( e^{(\mu x)^{\sigma}} - 1 \right)^\nu \right]^{-1}, \quad x>0,$$

respectively, where $\mu>0, \quad \sigma\nu>0$. $\mu$ is the scale parameter, $\sigma$ and $\nu$ are the shape parameters. Next figure shows possible shapes of the $f(x)$ and $F(x)$ for several values of the parameters.

```{r, echo=FALSE, message=F}
require(RelDists)
## 5^(1/3) = 1.71; 2^(1/3) = 1.26; 0.5^(1/2) = 0.71
curve(dOW(x, 1.71, 3, 0.7), from=0, to=3.5, ylim=c(0,1.6), col="red",
          ylab=expression(f[' '] (x)), xlab="x", las=1, lwd=2)
curve(dOW(x, 1.26, 3, 0.2), from=0, to=3.5, col="blue", 
      add=TRUE, lwd=2, lty=3) 
curve(dOW(x,0.71, 2, 1.3), from=0, to=3.5, col="darkgreen", 
      add=TRUE, lwd=2, lty=2)
curve(dOW(x,1.7,1,0.8), from=0, to=3.5, col="purple", 
      add=TRUE, lwd=2, lty=4)
legend("topright", legend=c(expression(paste(mu,
       " = ", 1.71,"      " ,sigma," = ",3,"   " ,nu," = ", 0.7)),
                            expression(paste(mu, 
       " = ", 1.26,"      " ,sigma," = ",3,"   " ,nu," = ", 0.2)),
                            expression(paste(mu,
       " = ", 0.71,"      " ,sigma," = ",2,"   " ,nu," = ", 1.3)),
                            expression(paste(mu,
       " = ", 1.7,"        " ,sigma," = ",1,"   " ,nu," = ", 0.8))),
       col=c("red","blue","darkgreen","purple"),lty=c(1,3,2,4), 
       bty="n", lwd=2)

## 5^(1/3) = 1.71; 2^(1/3) = 1.26; 0.5^(1/2) = 0.71
curve(pOW(x, 1.71, 3, 0.7), from=0, to=3.5, col="red",
      ylab=expression(F[' '] (x)), xlab="x", las=1, lwd=2)
curve(pOW(x, 1.26, 3, 0.2), from=0, to=3.5, col="blue", 
      add=TRUE, lwd=2, lty=3) 
curve(pOW(x,0.71, 2, 1.3), from=0, to=3.5, col="darkgreen", 
      add=TRUE, lwd=2, lty=2)
curve(pOW(x,1.7,1,0.8), from=0, to=3.5, col="purple", 
      add=TRUE, lwd=2, lty=4)
legend("bottomright", legend=c(expression(paste(mu, 
       " = ", 1.71,"      " ,sigma," = ",3,"   " ,nu," = ", 0.7)),
                            expression(paste(mu,
       " = ", 1.26,"      " ,sigma," = ",3,"   " ,nu," = ", 0.2)),
                            expression(paste(mu, 
       " = ", 0.71,"      " ,sigma," = ",2,"   " ,nu," = ", 1.3)),
                            expression(paste(mu,
       " = ", 1.7,"        " ,sigma," = ",1,"   " ,nu," = ", 0.8))),
       col=c("red","blue","darkgreen","purple"),lty=c(1,3,2,4), 
       bty="n", lwd=2)
```


The hf of the OW distribution is:

$$h(x) = \left( \frac{\sigma\nu}{x} \right) (\mu x)^\sigma e^{(\mu x)^\sigma} \left(e^{(\mu x)^{\sigma}}-1\right)^{\nu-1} \left[ 1 + \left(e^{(\mu x)^{\sigma}}-1\right)^\nu \right]^{-1}, x>0,$$
        
where the hf can take the following shapes:

+ Increasing if $\sigma>1$ and $\sigma\nu>1$.

+ Decreasing if $\sigma<1$ and $\sigma\nu<1$.

+ Unimodal shaped if either $\sigma<0$ and $\nu<0$ or $\sigma<1$ and $\sigma\nu\geq 1$.

+ Bathtub shaped if $\sigma>1$ and $\sigma\nu\geq 1$.
      
The figure shows possible shapes of the hf mentioned above: 

```{r, echo=FALSE}
par(mgp=c(3,0.7,0))
curve(hOW(x, 1/85, 9, 0.7), from=0, to=69, ylim=c(0,0.035), col="red",
      ylab=expression(h[' '] (x)), xlab="x", las=1, lwd=2, lty=1)
curve(hOW(x, 1/100, 0.5, 0.3), from=0, to=70, col="blue", 
      add=TRUE, lwd=2, lty=3) 
curve(hOW(x, 1/50, 1, 1), from=0, to=70, col="darkgreen", 
      add=TRUE, lwd=2, lty=2)
curve(hOW(x, 1/45, 8, 0.01), from=0, to=69, col="purple", 
      add=TRUE, lwd=2, lty=4)
curve(hOW(x, 1/75, -1.5, -0.1), from=0, to=70, col="black", 
      add=TRUE, lwd=2)
legend("topright", legend=c(expression(paste(mu, 
       " = ", 0.012,"     " ,sigma," = ",9,"        " ,nu," = ", 0.7)),
                            expression(paste(mu, 
       " = ", 0.01,"       " ,sigma," = ",0.5,"     " ,nu," = ", 0.3)),
                            expression(paste(mu, 
       " = ", 0.02,"       " ,sigma," = ",1,"        " ,nu," = ", 1)),
                            expression(paste(mu, 
       " = ", 0.0222,"  " ,sigma," = ",8,"         " ,nu," = ", 0.01)),
                            expression(paste(mu, 
       " = ", 0.0133,"  " ,sigma," = ",-1.5,"   " ,nu," = ", -0.1))), 
       col=c("red","blue","darkgreen","purple","black"),lty=c(1,3,2,4,1), 
       bty="n", lwd=2)
```

# Application examples

## Time to failure on electronic equipment

Cooray [-@Cooray2015] used the following data provided by Wang [-@Wang2000] in order to fit an OW distribution through maximum likelihood estimation (MLE):
 
5, 11, 21, 31, 46, 75, 98, 122, 145, 165, 195, 224, 245, 293, 321, 330, 350, 420.

The data above is the time to failure of an electronic device in hours. As an alternative to classical MLE, We used the function \verb|gamlss| to fit an only-intercept model in order to estimate parameters of OW distribution without covariates. Using our `initValuesOW_TTT()`, we can obtain an initial guess and the valid region.

```{r}
require(RelDists)

data("equipment")
my_initial_guess <- initValuesOW_TTT(formula = equipment ~ 1)

summary(my_initial_guess)
```

`initValuesOW_TTT()` function detected the Bathtub hazard shape, which corresponds to a convex-then-concave shape of total time on test (TTT) plot

```{r}
plot(my_initial_guess)
```

Therefore, we define the search region according to `initValuesOW_TTT()` outputs

```{r}
# Custom search region
myvalues <- list(sigma="all(sigma > 1)",
                 nu="all(nu < 1) & all(nu < 1)")
```

and we perform the fit using `gamlss()`

```{r, message=FALSE, warning=FALSE}
require(gamlss)
param_ee <- gamlss(equipment ~ 1, sigma.fo = ~ 1, nu.fo = ~ 1, 
                   sigma.start = 5, nu.start = 0.1,
                   control = gamlss.control(n.cyc=300, trace=FALSE),
                   family = myOW_region(family = OW(sigma.link='identity'),
                                        valid.values = myvalues,
                                        initVal = my_initial_guess))
summary(param_ee)
```
```{r, echo=FALSE, include=FALSE}
mu1 <- exp(coef(param_ee, what='mu'))
sigma1 <- coef(param_ee, what='sigma')
nu1 <- exp(coef(param_ee, what='nu'))

mymu1 <- formatC(mu1, format = "e", digits = 2)
mysigma1 <- round(sigma1, 4)
mynu1 <- round(nu1, 4)
```

In the following table we summarize the results and compare them with those gotten by Cooray [-@Cooray2015]

| Parameter | MLE [@Cooray2015]  |  GAMLSS     |
|:---------:|:------------------:|:-----------:|
| $\mu$     | 5.35e-03           | `r mymu1`   |
| $\sigma$  | 3.22388            | `r mysigma1`|
| $\nu$     | 0.28424            | `r mynu1`   |

## Mortality in mice exposed to radiation

Cooray [-@Cooray2006] used a dataset with 208 data points provided by Kimball [-@Kimball1960] with the ages at death in weeks for male mice exposed to 240r of gamma radiation. Again, we implement a workflow for parameter estimation with `myOW_region` and `gamlss` functions.

```{r}
# Do not forget to load 'RelDists' package
data("mice")
init_vals <- initValuesOW_TTT(formula = mice ~ 1)

summary(init_vals)
```

With `initValuesOW_TTT()` function we identified an increasing hazard shape, as well as was stated by Cooray [-@Cooray2006], because TTT plot is concave.

```{r}
plot(init_vals)
```

Hence, we implement the estimation procedure


```{r, message=FALSE, warning=FALSE}
# Custom search region
# Do not forget to load 'gamlss' library
param_mm <- gamlss(mice ~ 1, sigma.fo = ~ 1, nu.fo = ~ 1,
                   sigma.start = 5, nu.start = 0.1,
                   control = gamlss.control(n.cyc=300, trace=FALSE),
                   family = myOW_region(family = OW(sigma.link='identity'),
                                        valid.values = "auto",
                                        initVal = init_vals))
summary(param_mm)
```
```{r, echo=FALSE, include=FALSE}
mu2 <- exp(coef(param_mm, what='mu'))
sigma2 <- coef(param_mm, what='sigma')
nu2 <- exp(coef(param_mm, what='nu'))

mymu2 <- formatC(mu2, format = "e", digits = 2)
mysigma2 <- round(sigma2, 4)
mynu2 <- round(nu2, 4)
```

Then, we report the results and compare them with those in Cooray [-@Cooray2006]

| Parameter | MLE [@Cooray2006]  |  GAMLSS     |
|:---------:|:------------------:|:-----------:|
| $\mu$     | 7.61e-03           | `r mymu2`   |
| $\sigma$  | 6.2278             | `r mysigma2`|
| $\nu$     | 0.7495             | `r mynu2`   |

# References
