---
title: "OW distribution"
author: "Jaime Mosquera"
date: "`r Sys.Date()`"
bibliography: ../inst/REFERENCES.bib
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{OW distribution}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE}
library(RelDists)
library(EstimationTools)
library(gamlss)
```

# Odd Weibull distribution

This distribution was proposed by Cooray [-@Cooray2006]. The probability density function $f(x)$ and cumulative density function $F(x)$ are given by:

$$f(x) = \left( \frac{\sigma\nu}{x} \right) (\mu x)^\sigma e^{(\mu x)^\sigma} \left(e^{(\mu x)^{\sigma}}-1\right)^{\nu-1} \left[ 1 + \left(e^{(\mu x)^{\sigma}}-1\right)^\nu \right]^{-2},$$

and

$$F(x) = 1 - \left[ 1 + \left( e^{(\mu x)^{\sigma}} - 1 \right)^\nu \right]^{-1}, \quad x>0,$$

respectively, where $\mu>0, \quad \sigma\nu>0$. $\mu$ is the scale parameter, $\sigma$ and $\nu$ are the shape parameters. Next figure shows possible shapes of the $f(x)$ and $F(x)$ for several values of the parameters.

```{r, OWplot01, fig.alt="OW pdf and cdf", echo=FALSE, message=F}
old_par <- par(mfrow = c(1, 1)) # save previous graphical parameters

## 5^(1/3) = 1.71; 2^(1/3) = 1.26; 0.5^(1/2) = 0.71
paleta <- c("#00004A", "#00A4FF", "#F6F906", "#FF3300", "purple")
curve(dOW(x, 1.71, 3, 0.7), from=0, to=3.5, ylim=c(0,1.6), col=paleta[1],
          ylab=expression(f[' '] (x)), xlab="x", las=1, lwd=2)
curve(dOW(x, 1.26, 3, 0.2), from=0, to=3.5, col=paleta[2], 
      add=TRUE, lwd=2) 
curve(dOW(x,0.71, 2, 1.3), from=0, to=3.5, col=paleta[3], 
      add=TRUE, lwd=2)
curve(dOW(x,1.7,1,0.8), from=0, to=3.5, col=paleta[4], 
      add=TRUE, lwd=2)
legend("topright", legend=c(expression(paste(mu,
       " = ", 1.71,"      " ,sigma," = ",3,"   " ,nu," = ", 0.7)),
                            expression(paste(mu, 
       " = ", 1.26,"      " ,sigma," = ",3,"   " ,nu," = ", 0.2)),
                            expression(paste(mu,
       " = ", 0.71,"      " ,sigma," = ",2,"   " ,nu," = ", 1.3)),
                            expression(paste(mu,
       " = ", 1.7,"        " ,sigma," = ",1,"   " ,nu," = ", 0.8))),
       col=paleta[1:4], bty="n", lwd=2)

## 5^(1/3) = 1.71; 2^(1/3) = 1.26; 0.5^(1/2) = 0.71
curve(pOW(x, 1.71, 3, 0.7), from=0, to=3.5, col=paleta[1],
      ylab=expression(F[' '] (x)), xlab="x", las=1, lwd=2)
curve(pOW(x, 1.26, 3, 0.2), from=0, to=3.5, col=paleta[2], 
      add=TRUE, lwd=2) 
curve(pOW(x,0.71, 2, 1.3), from=0, to=3.5, col=paleta[3], 
      add=TRUE, lwd=2)
curve(pOW(x,1.7,1,0.8), from=0, to=3.5, col=paleta[4], 
      add=TRUE, lwd=2)
legend("bottomright", legend=c(expression(paste(mu, 
       " = ", 1.71,"      " ,sigma," = ",3,"   " ,nu," = ", 0.7)),
                            expression(paste(mu,
       " = ", 1.26,"      " ,sigma," = ",3,"   " ,nu," = ", 0.2)),
                            expression(paste(mu, 
       " = ", 0.71,"      " ,sigma," = ",2,"   " ,nu," = ", 1.3)),
                            expression(paste(mu,
       " = ", 1.7,"        " ,sigma," = ",1,"   " ,nu," = ", 0.8))),
       col=paleta[1:4], bty="n", lwd=2)

par(old_par) # restore previous graphical parameters
```


The hazard function (hf) of the OW distribution is:

$$h(x) = \left( \frac{\sigma\nu}{x} \right) (\mu x)^\sigma e^{(\mu x)^\sigma} \left(e^{(\mu x)^{\sigma}}-1\right)^{\nu-1} \left[ 1 + \left(e^{(\mu x)^{\sigma}}-1\right)^\nu \right]^{-1}, x>0,$$
        
where the hf can take the following shapes:

+ Increasing if $\sigma>1$ and $\sigma\nu>1$.

+ Decreasing if $\sigma<1$ and $\sigma\nu<1$.

+ Unimodal shaped if either $\sigma<0$ and $\nu<0$ or $\sigma<1$ and $\sigma\nu\geq 1$.

+ Bathtub shaped if $\sigma>1$ and $\sigma\nu\geq 1$.
      
The figure shows possible shapes of the hf mentioned above: 

```{r, OWplot02, fig.alt="OW hazard", echo=FALSE}
old_par <- par(mfrow = c(1, 1)) # save previous graphical parameters

par(mgp=c(3,0.7,0))
curve(hOW(x, 1/85, 9, 0.7), from=0, to=69, ylim=c(0,0.035), col=paleta[1],
      ylab=expression(h[' '] (x)), xlab="x", las=1, lwd=2)
curve(hOW(x, 1/100, 0.5, 0.3), from=0, to=70, col=paleta[2], 
      add=TRUE, lwd=2) 
curve(hOW(x, 1/50, 1, 1), from=0, to=70, col=paleta[3], 
      add=TRUE, lwd=2)
curve(hOW(x, 1/45, 8, 0.01), from=0, to=69, col=paleta[4], 
      add=TRUE, lwd=2)
curve(hOW(x, 1/75, -1.5, -0.1), from=0, to=70, col=paleta[5], 
      add=TRUE, lwd=2)
legend("topright", legend=c(expression(paste(mu, 
       " = ", 0.012,"     " ,sigma," = ",9,"        " ,nu," = ", 0.7)),
                            expression(paste(mu, 
       " = ", 0.01,"       " ,sigma," = ",0.5,"     " ,nu," = ", 0.3)),
                            expression(paste(mu, 
       " = ", 0.02,"       " ,sigma," = ",1,"        " ,nu," = ", 1)),
                            expression(paste(mu, 
       " = ", 0.0222,"  " ,sigma," = ",8,"         " ,nu," = ", 0.01)),
                            expression(paste(mu, 
       " = ", 0.0133,"  " ,sigma," = ",-1.5,"   " ,nu," = ", -0.1))), 
       col=paleta, bty="n", lwd=2)

par(old_par) # restore previous graphical parameters
```

The following figure illustrate the regions corresponding to the different hazard shapes:

```{r, OWplot03, fig.alt="OW hazard shapes", echo=FALSE, message=FALSE, warning=FALSE, fig.height=6}
old_par <- par(mfrow = c(1, 1)) # save previous graphical parameters

hyp <- function(x) 1/x

library(viridis)
col1 <- magma(18)[15]
col2 <- magma(18)[12]
col3 <- magma(18)[7]
col4 <- magma(18)[10]
col5 <- magma(18)[1]
col6 <- "blue"

par(mfrow=c(1,1), mar=c(5,5,1,1), mai=c(1,1,0.2,0.2), 
    mgp=c(3.6,1.1,0))
x <- seq(-3.5, 3.5, length.out = 500)
y <- hyp(x)
plot(x, y, type="l", xlim=c(-3,3), ylim=c(-3,3),
     xlab=expression(sigma), ylab=expression(nu), las=1)

# Increasing zone
x_inc <- x[which(x > 1)]
y_inc <- hyp(x_inc)
polygon(x = c(1, 1, x_inc, x_inc[length(x_inc)]), border = 1,
        y = c(3.5, 1, y_inc, 3.5), col=col1)

# Unimodal zone
x_unim1 <- x[which(x < 1 & x > 1/3.5)]
y_unim1 <- hyp(x_unim1)
polygon(x = c(x_unim1, 1, 1, 1/3.5), border = 1,
        y = c(y_unim1, 1, 3.5, 3.5), col=col2)

x_unim2 <- x[which(x < -1/3.5)]
y_unim2 <- hyp(x_unim2)
polygon(x = c(x_unim2, -1/3.5, -3.5, -3.5), border = 1,
        y = c(y_unim2, -3.5, -3.5, -1/3.5), col=col3)

# Bathtub zone
polygon(x = c(1, x_inc, 3.5, 3.5, 1), border = 1,
        y = c(1, y_inc, 1/3.5, 0, 0), col=col4)

# Decreasing zone
polygon(x = c(x_unim1, 1, 1, 0, 0), border = 1,
        y = c(y_unim1, 1, 0, 0, 3.5), col=col6)
polygon(x = c(x_unim2, -1/3.5, 0, 0, -3.5, -3.5),
        y = c(y_unim2, -3.5, -3.5, 0, 0, -1/3.5), col=col5)
box()
abline(h=0); abline(v=0)
legend("topleft", legend=c("Increasing", "Unimodal 1", 
                           "Unimodal 2", "Bathtub", 
                           "Decreasing 1", "Decreasing 2"),
       col=1, pch=21, pt.cex=2.5, pt.bg=c(col1,col2,col3,col4,col6,col5), 
       y.intersp=1.2)

par(old_par) # restore previous graphical parameters
```


# Application examples

<!-- In the examples below the following model is fitted -->

<!-- \begin{align} -->
<!-- y &\sim OW(\mu,\: \sigma,\: \nu) \\ -->
<!-- \log (\mu) &= \alpha_0 \\ -->
<!-- \sigma &= \beta_0 \\ -->
<!-- \log (\nu) &= \gamma_0 -->
<!-- \end{align} -->

## Time to failure on electronic equipment

Cooray [-@Cooray2015] used the following data provided by Wang [-@Wang2000] in order to fit an OW distribution through maximum likelihood estimation (MLE):
 
5, 11, 21, 31, 46, 75, 98, 122, 145, 165, 195, 224, 245, 293, 321, 330, 350, 420.

The data above is the time to failure of an electronic device in hours. As an alternative to classical MLE, We used the function \verb|gamlss| to fit an only-intercept model in order to estimate parameters of OW distribution without covariates. Using our `initValuesOW()`, we can obtain an initial guess and the valid region.

```{r}
data("equipment")
my_initial_guess <- initValuesOW(formula = equipment ~ 1)

summary(my_initial_guess)
```

`initValuesOW()` function detected the Bathtub hazard shape, which corresponds to a convex-then-concave shape of total time on test (TTT) plot

```{r OWplot04, fig.alt="OW Hazard shape",}
old_par <- par(mfrow = c(1, 1)) # save previous graphical parameters

par(mar = c(3.7, 3.7, 1, 10), mgp = c(2.5, 1, 0))
plot(my_initial_guess, las = 1)
legend.HazardShape(x = 1.07, y = 1.04, xpd = TRUE)

par(old_par) # restore previous graphical parameters
```

Therefore, we define the search region according to `initValuesOW()` outputs

```{r}
# Custom search region
myvalues <- list(sigma = "all(sigma > 1)",
                 nu = "all(nu < 1/sigma)")
```

and we perform the fit using `gamlss()`

```{r, message=FALSE, warning=FALSE}
# gamlss set up
con.out <-gamlss.control(n.cyc = 300, trace=TRUE)
myOW <- myOW_region(family = OW(sigma.link='identity'),
                    valid.values = myvalues, initVal = my_initial_guess)

param_ee <- gamlss(equipment ~ 1, sigma.fo = ~ 1, nu.fo = ~ 1, 
                   sigma.start = 5, nu.start = 0.1,
                   control = con.out, family = myOW)
summary(param_ee)
```
```{r, echo=FALSE, include=FALSE}
mu1 <- exp(coef(param_ee, what = 'mu'))
sigma1 <- coef(param_ee, what = 'sigma')
nu1 <- exp(coef(param_ee, what = 'nu'))

mymu1 <- formatC(mu1, format = "e", digits = 2)
mysigma1 <- round(sigma1, 4)
mynu1 <- round(nu1, 4)
```

In the following table we summarize the results and compare them with those gotten by Cooray [-@Cooray2015]

| Parameter | MLE [@Cooray2015]  |  GAMLSS     |
|:---------:|:------------------:|:-----------:|
| $\mu$     | 5.35e-03           | `r mymu1`   |
| $\sigma$  | 3.22388            | `r mysigma1`|
| $\nu$     | 0.28424            | `r mynu1`   |

## Mortality in mice exposed to radiation

Cooray [-@Cooray2006] used a dataset with 208 data points provided by Kimball [-@Kimball1960] with the ages at death in weeks for male mice exposed to 240r of gamma radiation. Again, we implement a workflow for parameter estimation with `myOW_region` and `gamlss` functions.

```{r}
# Do not forget to load 'RelDists' package
data("mice")
init_vals <- initValuesOW(formula = mice ~ 1)

summary(init_vals)
```

With `initValuesOW()` function we identified an increasing hazard shape, as well as was stated by Cooray [-@Cooray2006], because TTT plot is concave.

```{r OWplot05, fig.alt="OW empirical hazard shape",}
old_par <- par(mfrow = c(1, 1)) # save previous graphical parameters

par(mar = c(3.7, 3.7, 1, 10), mgp = c(2.5, 1, 0))
plot(init_vals, las = 1)
legend.HazardShape(x = 1.07, y = 1.04, xpd = TRUE)

par(old_par) # restore previous graphical parameters
```

Hence, we implement the estimation procedure


```{r, message=FALSE, warning=FALSE, message=FALSE}
# gamlss set up
myOW <- myOW_region(initVal = init_vals)

# Custom search region
# Do not forget to load 'gamlss' library
param_mm <- gamlss(mice ~ 1, sigma.fo = ~ 1, nu.fo = ~ 1,
                   sigma.start = 2, nu.start = 6,
                   control = con.out,
                   family = myOW)
summary(param_mm)
```
```{r, echo=FALSE, include=FALSE}
mu2 <- exp(coef(param_mm, what = 'mu'))
sigma2 <- coef(param_mm, what = 'sigma')
nu2 <- exp(coef(param_mm, what = 'nu'))

mymu2 <- formatC(mu2, format = "e", digits = 2)
mysigma2 <- round(exp(sigma2), 4)
mynu2 <- round(nu2, 4)
```

Then, we report the results and compare them with those in Cooray [-@Cooray2006]

| Parameter | MLE [@Cooray2006]  |  GAMLSS     |
|:---------:|:------------------:|:-----------:|
| $\mu$     | 7.61e-03           | `r mymu2`   |
| $\sigma$  | 6.2278             | `r mysigma2`|
| $\nu$     | 0.7495             | `r mynu2`   |

# References
