---
title: "OW distribution"
author: "Jaime Mosquera"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "##"
)
```

## Odd Weibull distribution

In this vignette we explore the Odd Weibull distribution.

The distribution was proposed by Cooray (2006).

### Probability density function 
The odd Weibull distribution (OW) was proposed by Cooray [-@Cooray2006]. It has three parameters. 

The pdf of the OW distributon is:

$$f(x)=\mu\theta\nu x^{\theta-1} e^{\mu x^{\theta}}
      (e^{\mu x^{\theta}}-1)^{\nu -1}[1+(e^{\mu      
      x^{\theta}}-1)^{\nu}]^{-2}, x>0,$$

where $\mu>0, \theta>0$ and $\nu>0$, the shape parameters being $\theta$ and $\nu$. 

The figure shows possible shapes of the pdf: 

```{r, echo=FALSE, message=F}
require(RelDists)
curve(dOW(x,5,3,0.7), from=0, to=3.5, ylim=c(0,2), col="red",ylab="Probability density function", las=1)
curve(dOW(x,2,3,0.2), col="blue", add=T) 
curve(dOW(x,0.5,2,1.3), col="darkgreen", add=T)
legend("topright", legend=c(expression(paste(mu, " = ", 5,"      " ,theta,"=",3,"   " ,nu," = ", 0.7)),
                              expression(paste(mu, " = ", 2,"      " ,theta,"=",3,"   " ,nu," = ", 0.2)),
                              expression(paste(mu, " = ", 0.5,"   " ,theta,"=",2,"   " ,nu," = ", 1.3))), 
       col=c("red","blue","darkgreen"),lty=c(1,1,1), bty="n")
```


### Cumulative density function 
The cdf of the OW distributon is:

$$F(x)=1-[1+(e^{\mu x^{\theta}}-1^\nu)]^{-1}, x>0.$$

The figure shows possible shapes of the cdf: 

```{r, echo=FALSE}
curve(pOW(x,5,3,0.7), from=0, to=3.5, ylim=c(0,1), col="red",ylab="Cumulative distribution function", las=1)
curve(pOW(x, 2,3,0.2), col="blue", add=T)
curve(pOW(x,0.5,2,1.3), col="darkgreen", add=T)
legend("bottomright", legend=c(expression(paste(mu, " = ", 5,"       " ,theta,"=",3,"    " ,nu," = ", 0.7)),
                              expression(paste(mu, " = ", 2,"       " ,theta,"=",3,"    " ,nu," = ", 0.2)),
                              expression(paste(mu, " = ", 0.5,"    " ,theta,"=",2,"    " ,nu," = ", 1.3))), 
       col=c("red","blue","darkgreen"),lty=c(1,1,1), bty="n")
```


### Hazard function
The hf of the OW distribution is:

$$h(x)=\mu\theta\nu x^{\theta-1} e^{\mu x^{\theta}}
      (e^{\mu x^{\theta}}-1)^{\nu -1}[1+(e^{\mu      
      x^{\theta}}-1)^{\nu}]^{-1}, x>0,$$
        
where the hf can take the following shapes:

+ Increasing if $\theta > 1$ and $\theta\nu > 1$.

+ Decreasing if $\theta < 1$ and $\theta\nu < 1$.

+ Unimodal shaped if either $\theta, \nu < 0$ or $\theta < 1$ and $\theta\nu \ge 1$.

+ Bathtub shaped if $\theta > 1$ and $\theta\nu \leq 1$.
      
The figure shows possible shapes of the hf mentioned above: 

```{r, echo=FALSE}
curve(hOW(x,4,3,0.7), from=0, to=2.5, ylim=c(0,3), col="red",ylab="Hazard function", las=1)
curve(hOW(x,0.5,0.8,0.7),   col="blue", add=T)
curve(hOW(x,0.5,0.5,3), col="darkgreen", add=T)
curve(hOW(x,0.8,3.5,0.2), col="purple", add=T)  
legend("top", legend=c(expression(paste(mu, " = ", 4,"       " ,theta,"=",3,"      " ,nu," = ", 0.7)),
                              expression(paste(mu, " = ", 0.5,"    " ,theta,"=",0.8,"   " ,nu," = ", 0.7)),
                              expression(paste(mu, " = ", 0.5,"    " ,theta,"=",0.5,"   " ,nu," = ", 3)), expression(paste(mu, " = ", 0.8,"    " ,theta,"=",3.5,"   " ,nu," = ", 0.2))), 
       col=c("red","blue","darkgreen","purple"),lty=c(1,1,1,1), bty="n")
```