---
title: "OW distribution"
author: ""
date: "`r Sys.Date()`"
bibliography: distributions.bib
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "##"
)
```

## Odd Weibull distribution

In this vignette we explore the Odd Weibull distribution.

The distribution was proposed by Cooray [-@Cooray2006].

### Probability density function 

$$f(x) = \left( \frac{\sigma\nu}{x} \right) (\mu x)^\sigma e^{(\mu x)^\sigma} \left(e^{(\mu x)^{\sigma}}-1\right)^{\nu-1} \left[ 1 + \left(e^{(\mu x)^{\sigma}}-1\right)^\nu \right]^{-2}, \quad x>0, \quad \mu>0, \quad \sigma\nu>0$$

where $\mu$ is the scale parameter, $\sigma$ and $\nu$ are the shape parameters.

The figure shows possible shapes of the pdf: 

```{r, echo=FALSE, message=F}
require(RelDists)
## 5^(1/3) = 1.71; 2^(1/3) = 1.26; 0.5^(1/2) = 0.71
curve(dOW(x, 1.71, 3, 0.7), from=0, to=3.5, ylim=c(0,1.6), col="red",
          ylab=expression(f[' '] (x)), xlab="x", las=1, lwd=2)
curve(dOW(x, 1.26, 3, 0.2), from=0, to=3.5, col="blue", 
      add=TRUE, lwd=2, lty=3) 
curve(dOW(x,0.71, 2, 1.3), from=0, to=3.5, col="darkgreen", 
      add=TRUE, lwd=2, lty=2)
curve(dOW(x,1.7,1,0.8), from=0, to=3.5, col="purple", 
      add=TRUE, lwd=2, lty=4)
legend("topright", legend=c(expression(paste(mu, " = ", 1.71,"      " ,sigma," = ",3,"   " ,nu," = ", 0.7)),
                            expression(paste(mu, " = ", 1.26,"      " ,sigma," = ",3,"   " ,nu," = ", 0.2)),
                            expression(paste(mu, " = ", 0.71,"      " ,sigma," = ",2,"   " ,nu," = ", 1.3)),
                            expression(paste(mu, " = ", 1.7,"        " ,sigma," = ",1,"   " ,nu," = ", 0.8))),
       col=c("red","blue","darkgreen","purple"),lty=c(1,3,2,4), 
       bty="n", lwd=2)
```


### Cumulative density function 
The cdf of the OW distributon is:

$$F(x) = 1 - \left[ 1 + \left( e^{(\mu x)^{\sigma}} - 1 \right)^\nu \right]^{-1}, x>0.$$

The figure shows possible shapes of the cdf: 

```{r, echo=FALSE}
## 5^(1/3) = 1.71; 2^(1/3) = 1.26; 0.5^(1/2) = 0.71
curve(pOW(x, 1.71, 3, 0.7), from=0, to=3.5, col="red",
      ylab=expression(F[' '] (x)), xlab="x", las=1, lwd=2)
curve(pOW(x, 1.26, 3, 0.2), from=0, to=3.5, col="blue", 
      add=TRUE, lwd=2, lty=3) 
curve(pOW(x,0.71, 2, 1.3), from=0, to=3.5, col="darkgreen", 
      add=TRUE, lwd=2, lty=2)
curve(pOW(x,1.7,1,0.8), from=0, to=3.5, col="purple", 
      add=TRUE, lwd=2, lty=4)
legend("bottomright", legend=c(expression(paste(mu, " = ", 1.71,"      " ,sigma," = ",3,"   " ,nu," = ", 0.7)),
                            expression(paste(mu, " = ", 1.26,"      " ,sigma," = ",3,"   " ,nu," = ", 0.2)),
                            expression(paste(mu, " = ", 0.71,"      " ,sigma," = ",2,"   " ,nu," = ", 1.3)),
                            expression(paste(mu, " = ", 1.7,"        " ,sigma," = ",1,"   " ,nu," = ", 0.8))),
       col=c("red","blue","darkgreen","purple"),lty=c(1,3,2,4), 
       bty="n", lwd=2)
```


### Hazard function (hf)
The hf of the OW distribution is:

$$h(x) = \left( \frac{\sigma\nu}{x} \right) (\mu x)^\sigma e^{(\mu x)^\sigma} \left(e^{(\mu x)^{\sigma}}-1\right)^{\nu-1} \left[ 1 + \left(e^{(\mu x)^{\sigma}}-1\right)^\nu \right]^{-1}, x>0,$$
        
where the hf can take the following shapes:

+ Increasing if $\sigma>1$ and $\sigma\nu>1$.

+ Decreasing if $\sigma<1$ and $\sigma\nu<1$.

+ Unimodal shaped if either $\sigma<0$ and $\nu<0$ or $\sigma<1$ and $\sigma\nu\geq 1$.

+ Bathtub shaped if $\sigma>1$ and $\sigma\nu\geq 1$.
      
The figure shows possible shapes of the hf mentioned above: 

```{r, echo=FALSE}
par(mgp=c(3,0.7,0))
curve(hOW(x, 1/85, 9, 0.7), from=0, to=69, ylim=c(0,0.035), col="red",
      ylab=expression(h[' '] (x)), xlab="x", las=1, lwd=2, lty=1)
curve(hOW(x, 1/100, 0.5, 0.3), from=0, to=70, col="blue", 
      add=TRUE, lwd=2, lty=3) 
curve(hOW(x, 1/50, 1, 1), from=0, to=70, col="darkgreen", 
      add=TRUE, lwd=2, lty=2)
curve(hOW(x, 1/45, 8, 0.01), from=0, to=69, col="purple", 
      add=TRUE, lwd=2, lty=4)
curve(hOW(x, 1/75, -1.5, -0.1), from=0, to=70, col="black", 
      add=TRUE, lwd=2)
legend("topright", legend=c(expression(paste(mu, " = ", 0.012,"     " ,sigma," = ",9,"        " ,nu," = ", 0.7)),
                            expression(paste(mu, " = ", 0.01,"       " ,sigma," = ",0.5,"     " ,nu," = ", 0.3)),
                            expression(paste(mu, " = ", 0.02,"       " ,sigma," = ",1,"        " ,nu," = ", 1)),
                            expression(paste(mu, " = ", 0.0222,"  " ,sigma," = ",8,"         " ,nu," = ", 0.01)),
                            expression(paste(mu, " = ", 0.0133,"  " ,sigma," = ",-1.5,"   " ,nu," = ", -0.1))), 
       col=c("red","blue","darkgreen","purple","black"),lty=c(1,3,2,4,1), 
       bty="n", lwd=2)
```

## References