---
title: "Distributions"
author: "Olga Usuga, Freddy Hernández, Catherine Rincón y Mauricio Barrientos"
date: "`r Sys.Date()`"
output: 
  pdf_document: 
    number_sections: yes
    toc: yes
vignette: >
  %\VignetteIndexEntry{hola}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
require(RelDists)
knitr::opts_chunk$set(echo = FALSE)
```

# Exponentiated Weibull distribution

## Probability density function 
The exponentiated Weibull distribution (EW) has three parameters. The pdf of the EW distribution is:

$$f(x)=\lambda \alpha \theta x^{\theta-1} \exp \left(-\alpha x^{\theta} \right)
      \left(1-\exp \left(-\alpha x^{\theta} \right) \right)^{\lambda-1}, \quad x>0,$$
      
where $\alpha$ is a scale parameter, $\theta$ and $\lambda> 0$ are shape parameters.

## Particular cases
When $\alpha$, $\theta$ and $\lambda$ parameters assume some values we can obtain known probability density functions, in the next table we present the cases.

|$\alpha$   |  $\lambda$  |   $\theta$  |     *Distribution*               |
|:----------|:----------: | :---------: |   ----------------------------   |
|1          |             |             |      Weibull                     | 
|           |             |        2    |      Generalized Rayleigh (GR)   |
|           |             |        1    |      Generalized Exponential(GE) |
|           |      1      |        1    |      Exponential                 |
|           |      1      |        2    |      Rayleigh                    |
            
The figure \@ref(fig:ew1) shows possible shapes for the pdf.

```{r ew1, fig.cap='Pdf shapes for EW distribution.', fig.asp=0.5, fig.width=4}
par(mar=c(2, 2.5, 0, 0), cex=0.5)
curve(dEW(x,5,0.5,10), from=0, to=2, ylim=c(0,3), col="red",
      ylab="Probability density function", las=1)
curve(dEW(x,0.7,0.5,0.15), col="green", add=T)
curve(dEW(x,1,3,1), col="purple", add=T)

legend("topright",
       legend=c(expression(paste(alpha, " = ", 5,"        " ,
                                 theta,"=",0.5,"     ",lambda," = ", 10)),
                expression(paste(alpha, " = ", 0.7,"     ",theta,"=",0.5,"     ",lambda," = ", 0.15)),
expression(paste(alpha, " = ", 1,"        " ,theta,"=",3,"        " ,lambda," = ", 1))), 
       col=c("red","green","purple"),lty=c(1,1,1), bty="n")
```      





# Flexible Weibull extension distribution

## Probability density function
The flexible Weibull extension (FWE) has two parameters. The pdf of the FWE distribution is:

$$f(x) = \left( \alpha+ \frac\beta{x^2} \right) \exp \left( \alpha x - {\frac\beta x} \right) \exp\left(-\exp \left(\alpha x - \frac{\beta}{x} \right) \right), \quad x > 0.$$  

The figure shows possible shapes of the pdf: 

```{r fig.cap='Pdf shapes for WE distribution.', fig.asp=0.5, fig.width=4}
par(mar=c(2, 2.5, 0, 0), cex=0.5)
curve(dFWE(x,0.75, 0.5), from=0, to=3, ylim=c(0,1.7), col="red",ylab="Probability density function", las=1)
curve(dFWE(x,2,3), col="blue", add=T)
curve(dFWE(x,0.75,1.3), col="green", add=T)
legend("topright", legend=c(expression(paste(alpha, " = ", 0.75,"   " ,beta,"=",0.5,"   " )),
  expression(paste(alpha, " = ", 2,"        " ,beta,"=",3,"   ")),
  expression(paste(alpha, " = ", 0.75,"   " ,beta,"=",1.3,"   "))), 
       col=c("red","blue","green"),lty=c(1,1,1), bty="n")
```

## Cumulative density function 
The cdf of the FWE distributon is:

$$F(x) = 1 - \exp \left( -\exp \left( \alpha x - \frac{\beta}{x} \right) \right), x > 0.$$

The figure shows possible shapes of the cdf: 

```{r fig.cap='Cdf shapes for WE distribution.', fig.asp=0.5, fig.width=4}
par(mar=c(2, 2.5, 0, 0), cex=0.5)
curve(pFWE(x,0.75, 0.5), from=0, to=3,  col="red",ylab="Cumulative distribution function", las=1)
curve(pFWE(x,2,3), col="blue", add=T)
curve(pFWE(x,0.75,1.3), col="green", add=T)
 
legend("bottomright", legend=c(expression(paste(alpha, " = ", 0.75,"   " ,beta,"=",0.5,"   " )),
  expression(paste(alpha, " = ", 2,"        " ,beta,"=",3,"   ")),
  expression(paste(alpha, " = ", 0.75,"   " ,beta,"=",1.3,"   "))), 
       col=c("red","blue","green"),lty=c(1,1,1), bty="n")
```

## Hazard function
The hf of the FWE distribution is:

$$h(x) =  \left( \alpha+ \frac\beta{x^2} \right) e^ {\alpha x - {\frac\beta x}}, \quad x > 0,$$

where $\alpha\beta > 0$.

The figure shows possible shapes of the hf:

```{r fig.cap='Hazard shapes for EW distribution.', fig.asp=0.5, fig.width=4}
par(mar=c(2, 2.5, 0, 0), cex=0.5)
curve(hFWE(x,alpha=0.7,   beta=1.3), from=0, to=3, ylim=c(0,3), col="red",ylab="Hazard function", las=1)
curve(hFWE(x,alpha=1.5,   beta=0.6), col="blue", add=T)
curve(hFWE(x,alpha=0.6,   beta=0.3), col="green", add=T)

legend("top", legend=c(expression(paste(alpha, " = ", 0.7,"     " ,beta,"=",1.3,"   " )),
  expression(paste(alpha, " = ", 1.5,"     " ,beta,"=",0.6,"   ")),
  expression(paste(alpha, " = ", 0.6,"     " ,beta,"=",0.3,"   "))), 
       col=c("red","blue","green"),lty=c(1,1,1), bty="n")
```

            