---
title: "Distributions"
author: "Olga Usuga, Freddy Hernández, Catherine Rincón y Mauricio Barrientos"
date: "`r Sys.Date()`"
output: 
  pdf_document: 
    number_sections: yes
    toc: yes
vignette: >
  %\VignetteIndexEntry{hola}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
require(RelDists)
knitr::opts_chunk$set(echo = FALSE)
```

# Exponentiated Weibull distribution

## Probability density function 
The exponentiated Weibull distribution (EW) has three parameters. The pdf of the EW distribution is:

$$f(x)=\lambda\alpha\theta x^{\theta-1} e^{-\alpha x^{\theta}}
      (1-e^{-\alpha x^{\theta}})^{\lambda-1}, \quad x>0,$$
      
where $\alpha$ is a scale parameter, $\theta$ and $\lambda> 0$ are shape parameters.

## Particular cases
When $\alpha$, $\theta$ and $\lambda$ parameters assume some values we can obtain known probability density functions, in the next table we present the cases.

|$\alpha$   |  $\lambda$  |   $\theta$  |     *Distribution*               |
|:----------|:----------: | :---------: |   ----------------------------   |
|1          |             |             |      Weibull                     | 
|           |             |        2    |      Generalized Rayleigh (GR)   |
|           |             |        1    |      Generalized Exponential(GE) |
|           |      1      |        1    |      Exponential                 |
|           |      1      |        2    |      Rayleigh                    |
            
The figure shows possible shapes for the pdf.

```{r fig.cap='Pdf shapes for EW distribution.', fig.asp=0.7, fig.width=5}
curve(dEW(x,5,0.5,10), from=0, to=2, ylim=c(0,3), col="red",
      ylab="Probability density function", las=1)
curve(dEW(x,0.7,0.5,0.15), col="green", add=T)
curve(dEW(x,1,3,1), col="purple", add=T)

legend("topright",
       legend=c(expression(paste(alpha, " = ", 5,"        " ,
                                 theta,"=",0.5,"     ",lambda," = ", 10)),
                expression(paste(alpha, " = ", 0.7,"     ",theta,"=",0.5,"     ",lambda," = ", 0.15)),
expression(paste(alpha, " = ", 1,"        " ,theta,"=",3,"        " ,lambda," = ", 1))), 
       col=c("red","green","purple"),lty=c(1,1,1), bty="n")
```      

            